Natural sources of uncertainty and their implications

### 2017-04-30 book_chapter/paper.tex
finished first write up of ref games section

The picture sketched in the last paragraph neglects one crucial thing: learning from observation. 
For rationality, we need to consider a rational agent's ability to learn from previous experience and to generalize from prior experience. 
For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice (\mf{insert refs for iterated learning}).
But even if every language learner in history had been perfectly rational, it is unlikely that anyone ever succeeded in reducing all uncertainty about language use in context and say with confidence: this is the threshold that truly separates heaps from non-heaps (\mf{ref / footnote Williamson?}).

 Which strategies for generalization are best depends on the stochastic properties of the environment in which learning and decision making takes place. The problem for language is that it what works best for on agent depends on what the other agent does. Communication problems are games in the formal sense of delivering payoffs only as a function of what all players do. Consequently, the problem of how best to adapt behavior from past experience in language games also depends on how other agents categorize, generalize and behave. Nothing gets off the ground with hyper-rationality alone. There needs to be learning, evolution and circumstances. As with sim-max games above, biases and other quirks of general cognition are likely only helpful in solving this massive coordination problem. Maybe our world is a place where particular learning biases are needed for any regular and shared communicative practice to get off the ground.

\begin{itemize}
\item uncertainty
\end{itemize}

What about meaning and vagueness in a universe of multiple language games and heavy uncertainty? We could look for meaning at several levels. Firstly, as before, there is the actual behavior of a single agent in each actualized language game. Like with sim-max games, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker). Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used, conditional on a given context. Either one of these two levels, the agent's behavior or linguistic generalizations, could be the foundation for a notion of emerging abstract meaning as constituted by the practices of a population at a particular time. The question is: for which of these three options of situating meaning would we like to have a theory of so that being vague would be disastrous for that theory?

>>>>

The picture sketched in the last paragraph neglects learning from observation and generalization across contexts. 
For rationality, we need to consider a rational agent's ability to learn from previous experience and to generalize from  the past to the future. 
For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice \parencite[\emph{e.g.}][]{KirbyGriffith2014:Iterated-Learni}.
But even if every language learner in history had been perfectly rational, it is extremely unlikely that anyone ever succeeded in reducing all uncertainty about language use and say with confidence: this is the threshold that truly separates heaps from non-heaps no matter what. 
For one, there may simply not be enough time in the lifespan of a rational learner to observe enough instances of language use for certainty about crisp denotations. 
For another, what needs to be learned likely changes over time. 
The optimal use for gradable properties hinges crucially on the prior over states \parencite[\emph{e.g.}][]{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie,QingFranke2014:Gradable-Adject,LassiterGoodman2015:Adjectival-vagu}. 
If the prior captures real world attributes, such as the probability of an individual's body height, these attributes will not only be hard to know for certain, but they will also vary over time. 
They may also depend on location: do we use our language exclusively in this village or do we travel beyond? 
If so, the priors that determine the goodness of a convention for one agent need not be the same as those for another. 
Going beyond the priors for property degrees, the probability of seeing a particular language game, e.g.~a particular context $c$ in some kind of reference game, is also crucial for determining globally efficient linguistic behavior. 
But here it is even more inconceivable how a rational learner could acquire a fairly confident estimate over how often a particular language game will be played. 
The list of sources of natural and seemingly insurmountable uncertainty about the parameters necessary to determine locally or globally optimal strategies is much longer than hinted at here. 
Still, it should be clear enough that a finite existence, a sparse learning input and the inherent fuzziness of the linguistic context itself (its temporal, local and social variability) is not a faul of rationality or the mechanism of selection of successful linguistic practices.

What about meaning and vagueness in a universe of multiple language games and heavy uncertainty? 
We could look for meaning at several levels. 
Firstly, as before, there is the actual behavior of a single agent in each actualized language game. 
As mentioned above, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker).
Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used, conditional on a given context. 
Notice that a single agent's rational beliefs about linguistic practices or linguistics meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
Either one of these two levels, the agent's behavior or linguistic generalizations, could be the foundation for a notion of emerging abstract meaning as constituted by the practices of a population at a particular time.
If both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp? 
Maybe there is a reason.
The main question to be answered is even more general: given a universe of (rational) language users who learn from finite experience in a changing and stochastically complex arrangement of multiple language games, is a notion of linguistic meaning a theoretically useful and empirically explanatory concept in this picture in such a way that if that notion entailed vague meanings the whole view would have disastrously little explanatory bite?

Taken together, when we look at the picture of multiple language games side by side, the question of what rational or optimal behavior is in any one of them, or the question of why human language users do not more closely approximate rational or optimal local behavior more closely seem entirely misplaced.
What seems more pressing is to answer how the interaction of individual learning, individual local optimization and general pressure for successful communication at a cross-situational level lead to the shaping of linguistic practices in an open-ended landscape of language games.
How to attain locally sub-optimal play in a single language game is not a theoretical puzzle we should take serious, if what really matters is a global multi-game perspective.
Adopting this perspective does not \emph{solve} the problem of how vagueness is compatible with rationality. 
But it suggest that it might be an inflated issue that we need not or should not worry about, because it will not matter for a more general and pressing issue in a more realistic and encompassing picture of language, its use and evolution. The following section will expand on this.
